---
title: "CSLS Lake Metrics"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Lake Metrics2}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSlevels)
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
library(extrafont)
library(lubridate)
library(NISTunits)
library(stringr)
library(reshape2)

# Parameters
text_size         <- 12
lakes             <- c("Pleasant", "Long", "Plainfield")

# Data
hist_levels       <- CSLSlevels::hist_levels
hist_levels_short <- hist_levels %>%
                     filter(year(.data$date) >= 1981,
                            year(.data$date) <= 2018)
hist_metrics       <- calculate_metrics(hist_levels_short)
hist_metrics_short <- CSLSlevels::hist_metrics_short %>%
                      filter(.data$lake != "Devils") %>%
                      group_by(.data$lake, .data$metric, .data$variable) %>%
                      summarise(median = median(.data$value),
                                sd = sd(.data$value)) %>%
                      ungroup()
hist_metrics_long  <- CSLSlevels::hist_metrics_long %>%
                      filter(.data$lake != "Devils") %>%
                      group_by(.data$lake, .data$metric, .data$variable) %>%
                      summarise(median = median(.data$value),
                                sd = sd(.data$value)) %>%
                      ungroup()

# Functions
create_table <- function(df, 
                         metric_name, 
                         variable_breaks, 
                         variable_labels,
                         convert_units = "",
                         round_precision = NA) {
  table_df <- df %>% 
              filter(.data$metric == metric_name) %>%
              select(.data$lake, .data$variable, .data$median, .data$sd)
  
  if (convert_units == "meterTOft") {
    table_df <- table_df %>% mutate_at(c("median", "sd"), NISTmeterTOft)
  }
  if (!is.na(round_precision)) {
    table_df$median <- round((1/round_precision)*table_df$median)/(1/round_precision)
    table_df$sd     <- round((1/round_precision)*table_df$sd)/(1/round_precision)
  }
  if (variable_breaks == "sort") {
    variable_breaks <- sort(as.numeric(unique(table_df$variable)))
  }
  if (variable_labels == "breaks") {
    variable_labels <- variable_breaks
  } else if (variable_labels == "percent") {
    variable_labels <- sprintf("%d%%", variable_breaks)
  }
  table_df$variable  <- factor(table_df$variable,
                               levels = variable_breaks,
                               labels = variable_labels)
  
  table_df$text <- sprintf("%s \U00B1 %s", table_df$median, table_df$sd)

  table_df <- dcast(table_df, lake~variable, value.var = "text")
  colnames(table_df) <- c("Lake", variable_labels)
  
  return(table_df)
}
```

<br>

# Historical Lake Levels

Estimated lake levels from 1938-2019 for each study lake (black line) with the
mean value of all estimations +/- one standard deviation shown in dark grey
(captures 68.2% of estimated values), the mean value of all estimations +/- two
standard deviations shown in light grey (captures 95% of estimated values), and
observations shown as points.

```{r levels01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=5}
plot_levels(hist_levels, 
            legend_pos = "top", 
            show_obs_type = TRUE, 
            color_vals = c("black", "#1B9E77", "#7570B3", "#D95F02"), 
            line_size = 0.7, 
            point_size = 1.5, 
            npretty_breaks = 5,  
            convert_units = "meterTOft", 
            grid_off = FALSE, 
            show_uncertainty = TRUE,
            text_size = text_size,
            lakes = c("Pleasant", "Long", "Plainfield"),
            force_range = FALSE)
```

<br>


```{r levels02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10.5, fig.height=6.5}
plot_levels(hist_levels_short, 
            legend_pos = "right", 
            show_obs_type = TRUE, 
            color_vals = c("black", "#1B9E77", "#7570B3", "#D95F02"), 
            line_size = 0.7, 
            point_size = 2.5, 
            npretty_breaks = 5,  
            convert_units = "meterTOft", 
            grid_off = FALSE, 
            show_uncertainty = TRUE,
            text_size = 24,
            lakes = c("Pleasant", "Long", "Plainfield"),
            force_range = FALSE)
```
```{r levels03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10.5, fig.height=6.5}
#7030A0
plot_levels(hist_levels_short, 
            legend_pos = "right", 
            color_vals = c("black", "grey40"), 
            line_size = 0.7, 
            point_size = 2.5, 
            npretty_breaks = 5,  
            convert_units = "meterTOft", 
            grid_off = FALSE, 
            show_uncertainty = FALSE,
            text_size = 24,
            lakes = c("Pleasant", "Long", "Plainfield"),
            force_range = FALSE)
```

```{r levels04, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10.5, fig.height=6.5}
plot_levels(hist_levels_short, 
            legend_pos = "right", 
            color_vals = c("black", "grey40"),
            all_probs = TRUE, 
            hline_color = c("#7030A0"),
            line_size = 0.7, 
            point_size = 2.5, 
            npretty_breaks = 5,  
            convert_units = "meterTOft", 
            grid_off = FALSE, 
            show_uncertainty = FALSE,
            text_size = 24,
            lakes = c("Pleasant", "Long", "Plainfield"),
            force_range = FALSE)
```

```{r levels05, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10.5, fig.height=6.5}

text_size <- 24
point_size = 2.5
line_size = 0.7
hline_color = c("#7030A0")
color_vals = c("black", "grey40")
legend_pos = "right"

# Filter to desired lakes
df <- hist_levels_short %>% 
      filter(.data$lake %in% c("Pleasant", "Long", "Plainfield"))
df <- df %>%
      mutate_at(c("level", "level_obs", "swims_obs", "usgs_obs",
                  "airphoto_obs", "mean", "sd", "p90", "p10","min",
                  "max", "upper", "lower"),
                NISTmeterTOft)
ytitle <- "Lake Elevation (ft)"
break_vals    <- c("Estimated", "Observation")
linetype_vals <- c("solid", "blank")
shape_vals    <- c(NA, 16)

exceeds <- calculate_exceedances(df, 50)
line_df <- filter(exceeds, variable == "50") %>%
           select(.data$lake, .data$value)
rect_df <- line_df %>%
           mutate(ymax = .data$value + 1,
                  ymin = .data$value - 1) %>%
           select("lake", "ymax", "ymin")

# Determine major/minor date breaks
date_min <- sprintf("%d-02-01", floor(year(min(df$date))/10)*10)
date_max <- sprintf("%d-02-01", ceiling(year(max(df$date))/10)*10)
if (max(year(df$date)) - min(year(df$date)) > 40) {
  breaks <- seq(as_datetime(date_min), as_datetime(date_max), "20 years")
} else {
  breaks <- seq(as_datetime(date_min), as_datetime(date_max), "10 years")
}
date_limits  <- c(as_datetime(date_min), as_datetime(date_max) - months(1))

plot_obj <- ggplot(data = df) +
            geom_line(aes(x = .data$date,
                          y = .data$level,
                          color = "Estimated"),
                      size = line_size) +
           geom_point(aes(x = .data$date,
                          y = .data$level_obs,
                          color = "Observation"),
                      size = point_size) +
           facet_wrap(~lake, scales = "free_y", ncol = 1) +
            geom_rect(data = rect_df,
                      mapping = aes(xmin = as_datetime(date_limits[1]),
                                    xmax = as_datetime(date_limits[2]),
                                    ymin = .data$ymin,
                                    ymax = .data$ymax),
                      alpha = 0.25,
                      fill = hline_color,
                      color = NA) +
            geom_hline(data = line_df,
                       aes(yintercept = .data$value),
                       color = hline_color,
                       linetype = "solid",
                       size = line_size) + 
            scale_y_continuous(breaks = scales::pretty_breaks(5)) +
            labs(x = "",
                 y = ytitle,
                 color = "",
                 title = "") +
            scale_color_manual(breaks = break_vals,
                               values = color_vals,
                               guide = guide_legend(override.aes = list(
                                 linetype = linetype_vals,
                                 shape = shape_vals))) +
            scale_x_datetime(breaks = breaks,
                             date_labels = "%Y",
                             expand = c(0,0),
                             limits = date_limits) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = text_size),
                  plot.title = element_text(hjust = 0.5),
                  legend.position = legend_pos,
                  legend.background = element_blank())
plot_obj
```

```{r levels06, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10.5, fig.height=6.5}

text_size <- 24
point_size = 2.5
line_size = 0.7
hline_color = c("#7030A0")
color_vals = c("black", "grey40")
legend_pos = "right"

# Filter to desired lakes
df <- hist_levels_short %>% 
      filter(.data$lake %in% c("Pleasant", "Long", "Plainfield"))
df <- df %>%
      mutate_at(c("level", "level_obs", "swims_obs", "usgs_obs",
                  "airphoto_obs", "mean", "sd", "p90", "p10","min",
                  "max", "upper", "lower"),
                NISTmeterTOft)
ytitle <- "Lake Elevation (ft)"
break_vals    <- c("Estimated", "Observation")
linetype_vals <- c("solid", "blank")
shape_vals    <- c(NA, 16)

exceeds <- calculate_exceedances(df, 10)
line_df <- exceeds %>%
           select(.data$lake, .data$value)
ribbon_df <- merge(df, line_df, by = "lake") %>%
             group_by(.data$lake, .data$date) %>%
             summarise(ymin = min(.data$level, .data$value),
                       ymax = .data$level) %>%
             ungroup()

# Determine major/minor date breaks
date_min <- sprintf("%d-02-01", floor(year(min(df$date))/10)*10)
date_max <- sprintf("%d-02-01", ceiling(year(max(df$date))/10)*10)
if (max(year(df$date)) - min(year(df$date)) > 40) {
  breaks <- seq(as_datetime(date_min), as_datetime(date_max), "20 years")
} else {
  breaks <- seq(as_datetime(date_min), as_datetime(date_max), "10 years")
}
date_limits  <- c(as_datetime(date_min), as_datetime(date_max) - months(1))

plot_obj <- ggplot(data = df) +
            geom_line(aes(x = .data$date,
                          y = .data$level,
                          color = "Estimated"),
                      size = line_size) +
           geom_point(aes(x = .data$date,
                          y = .data$level_obs,
                          color = "Observation"),
                      size = point_size) +
           facet_wrap(~lake, scales = "free_y", ncol = 1) +
            geom_ribbon(data = ribbon_df,
                       aes(x = .data$date,
                           ymin = .data$ymin,
                           ymax = .data$ymax),
                      alpha = 0.25,
                      fill = hline_color,
                      color = NA) +
            geom_hline(data = line_df,
                       aes(yintercept = .data$value),
                       color = hline_color,
                       linetype = "solid",
                       size = line_size) + 
            scale_y_continuous(breaks = scales::pretty_breaks(5)) +
            labs(x = "",
                 y = ytitle,
                 color = "",
                 title = "") +
            scale_color_manual(breaks = break_vals,
                               values = color_vals,
                               guide = guide_legend(override.aes = list(
                                 linetype = linetype_vals,
                                 shape = shape_vals))) +
            scale_x_datetime(breaks = breaks,
                             date_labels = "%Y",
                             expand = c(0,0),
                             limits = date_limits) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = text_size),
                  plot.title = element_text(hjust = 0.5),
                  legend.position = legend_pos,
                  legend.background = element_blank())
plot_obj
```

```{r levels07, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=10.5, fig.height=6.5}

text_size <- 24
point_size = 2.5
line_size = 0.7
hline_color = c("#7030A0")
color_vals = c("black", "grey40")
legend_pos = "right"

# Filter to desired lakes
df <- hist_levels_short %>% 
      filter(.data$lake %in% c("Pleasant", "Long", "Plainfield"))
df <- df %>%
      mutate_at(c("level", "level_obs", "swims_obs", "usgs_obs",
                  "airphoto_obs", "mean", "sd", "p90", "p10","min",
                  "max", "upper", "lower"),
                NISTmeterTOft)
ytitle <- "Lake Elevation (ft)"
break_vals    <- c("Estimated", "Observation")
linetype_vals <- c("solid", "blank")
shape_vals    <- c(NA, 16)

exceeds <- calculate_exceedances(df, 90)
line_df <- exceeds %>%
           select(.data$lake, .data$value)
ribbon_df <- merge(df, line_df, by = "lake") %>%
             group_by(.data$lake, .data$date) %>%
             summarise(ymax = max(.data$level, .data$value),
                       ymin = .data$level) %>%
             ungroup()

# Determine major/minor date breaks
date_min <- sprintf("%d-02-01", floor(year(min(df$date))/10)*10)
date_max <- sprintf("%d-02-01", ceiling(year(max(df$date))/10)*10)
if (max(year(df$date)) - min(year(df$date)) > 40) {
  breaks <- seq(as_datetime(date_min), as_datetime(date_max), "20 years")
} else {
  breaks <- seq(as_datetime(date_min), as_datetime(date_max), "10 years")
}
date_limits  <- c(as_datetime(date_min), as_datetime(date_max) - months(1))

plot_obj <- ggplot(data = df) +
            geom_line(aes(x = .data$date,
                          y = .data$level,
                          color = "Estimated"),
                      size = line_size) +
           geom_point(aes(x = .data$date,
                          y = .data$level_obs,
                          color = "Observation"),
                      size = point_size) +
           facet_wrap(~lake, scales = "free_y", ncol = 1) +
            geom_ribbon(data = ribbon_df,
                       aes(x = .data$date,
                           ymin = .data$ymin,
                           ymax = .data$ymax),
                      alpha = 0.25,
                      fill = hline_color,
                      color = NA) +
            geom_hline(data = line_df,
                       aes(yintercept = .data$value),
                       color = hline_color,
                       linetype = "solid",
                       size = line_size) + 
            scale_y_continuous(breaks = scales::pretty_breaks(5)) +
            labs(x = "",
                 y = ytitle,
                 color = "",
                 title = "") +
            scale_color_manual(breaks = break_vals,
                               values = color_vals,
                               guide = guide_legend(override.aes = list(
                                 linetype = linetype_vals,
                                 shape = shape_vals))) +
            scale_x_datetime(breaks = breaks,
                             date_labels = "%Y",
                             expand = c(0,0),
                             limits = date_limits) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = text_size),
                  plot.title = element_text(hjust = 0.5),
                  legend.position = legend_pos,
                  legend.background = element_blank())
plot_obj
```

```{r duration, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=8.5, fig.height=6.5}
plot_duration(hist_levels_short, 
              probs = c(10, 90), 
              show_median = FALSE, 
              bin_size = 1, 
              text_size = 24,
              max_months = 12) + 
  facet_grid(lake~variable, switch = "y") +
  theme(panel.spacing.x = unit(6, "mm"),
        panel.spacing.y = unit(7, "mm"))
```


<br>

```{r rate_of_change, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=8.5, fig.height=6.5}
plot_rate_change(hist_levels_short, 
                 text_size = 24,
                 density_color = "#7030A0",
                 show_median = FALSE) + 
  facet_grid(lake~variable, switch = "y") +
  theme(panel.spacing.x = unit(6, "mm"),
        panel.spacing.y = unit(7, "mm"))
```

<br>

```{r timing, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=8.5, fig.height=6.5}
plot_df <- hist_metrics_short %>%
           filter(.data$metric == "exceedance_range") %>%
           mutate(time = str_sub(.data$variable, 1, 1),
                  range = str_sub(.data$variable, 3, -1),
                  median = NISTunits::NISTmeterTOft(.data$median),
                  sd = NISTunits::NISTmeterTOft(.data$sd)) %>%
           mutate(range = str_replace(.data$range, "10_90", "10%-90%")) %>%
           mutate(range = str_replace(.data$range, "25_75", "25%-75%")) %>%
           mutate(text = sprintf("%s \U00B1 %s", .data$median, .data$sd)) %>%
           # filter(.data$range != "25%-75%") %>%
           select(.data$lake, .data$time, .data$range, .data$median, .data$sd)
plot_df <- dcast(plot_df, lake+range~time, value.var = "median")


ggplot(plot_df) +
  geom_point(aes(x = a, y = m, shape = lake, color = range), 
             # color = "black",
             size = 4) +
  geom_abline(slope = 1, intercept = 0) +
  labs(x = "Annual Range (ft)",
       y = "Monthly Range (ft)",
       shape = "") +
  theme_bw() +
  scale_color_manual(name = "",
                     breaks = c("10%-90%", "25%-75%"),
                     values = c("black", "grey70")) +
  theme(text = element_text(family = "Segoe UI Semilight",
                            size = 24),
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        legend.box = "vertical",
        legend.margin = margin(0,0,0,0))
```

```{r magnitude_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=5, fig.height=5,}
plot_magnitude(hist_levels_short, text_size = text_size)
```

<br>

```{r median_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_level"
metric_title    <- "**Median Lake Level (ft)**"
variable_breaks <- "sort"
variable_labels <- c("Annual","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r cv_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_level"
metric_title    <- "**CV of Lake Levels (%)**"
variable_breaks <- "sort"
variable_labels <- c("Annual","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r frequency, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=4}
plot_frequency(hist_levels_short, probs = c(10, 50, 90), text_size = text_size)
```

<br>

```{r exceedance_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6.5}
metric_name     <- "exceedance_level"
metric_title    <- "**Exceedance Probability Lake Levels (ft)**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

## 2. Frequency

```{r num_exceedances_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "num_dur_decade"
metric_title    <- "**Exceedances per Decade**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

```{r depart_median_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "depart_median"
metric_title    <- "**Probability within 1 ft of median (%)**"
variable_breaks <- c("0.3m", "-0.3m")
variable_labels <- c("+1ft", "-1ft")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

## 3. Duration


<br>

```{r median_dur_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_dur"
metric_title    <- "**Median Duration (months)**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_long, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r cv_dur_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_dur"
metric_title    <- "**CV of Duration (%)**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_long, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

## 4. Rate of change



<br>

```{r median_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_rise_rate"
metric_title    <- "**Median Rise Rate (ft/time period)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r cv_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_rise_rate"
metric_title    <- "**CV of Rise Rate (%)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r median_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_fall_rate"
metric_title    <- "**Median Fall Rate (ft/time period)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r cv_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_fall_rate"
metric_title    <- "**CV of Fall Rate (%)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

## 5. Timing



<br>

```{r exceedance_range_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "exceedance_range"
metric_title    <- "**Exceedance Probability Range (ft)**"
variable_breaks <- c("a_10_90","a_25_75","m_10_90","m_25_75")
variable_labels <- c("Annual 10%-90%", "Annual 25%-75%", 
                     "Monthly 10%-90%", "Monthly 25%-75%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
