---
title: "CSLS Lake Metrics - Sensitivity Analysis"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Lake Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSlevels)
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
library(extrafont)
library(lubridate)
library(NISTunits)

text_size              <- 12
convert_to_depth       <- TRUE
lakes                  <- c("Pleasant", "Long", "Plainfield")
csls_levels            <- CSLSlevels::csls_levels
csls_levels            <- csls_levels %>%
                          filter((.data$lake %in% lakes &
                                 year(.data$date) >= 1981 &
                                 year(.data$date) <= 2018) |
                                   (.data$lake == "Devils" &
                                 year(.data$date) >= 1981 &
                                 year(.data$date) <= 2001))

# Lake bottoms for CSLS lakes pulled from lake_raster
# Devil's Lake guesstimated based on "lake depth" in Wikipedia, the most reliable
# of sources.
monte_carlo_metrics    <- CSLSlevels::monte_carlo_metrics
if (convert_to_depth){
  lake_bottom <- data.frame(lake = c("Pleasant", "Long", "Plainfield", "Devils"),
                            bottom = c(291.1426, 332.8622, 332.0755,
                                       NISTftTOmeter(964-47)))
  monte_carlo_metrics <- merge(monte_carlo_metrics, 
                               lake_bottom, 
                               by = "lake", 
                               all.x = TRUE)
  monte_carlo_metrics <- monte_carlo_metrics %>%
                         mutate(sim = ifelse(.data$metric %in% c("median_level",
                                                                 "exceedance_level"),
                                             .data$sim - .data$bottom,
                                             .data$sim),
                                obs = ifelse(.data$metric %in% c("median_level",
                                                                 "exceedance_level"),
                                             .data$obs - .data$bottom,
                                             .data$obs))
  monte_carlo_metrics$bottom <- NULL
}

calculate_fit <- function(summary){
  fit <- summary %>%
         group_by(.data$lake, .data$metric, .data$variable, .data$nyear) %>%
         summarise(PBIAS = 100*abs(mean((.data$sim - .data$obs)/.data$obs, 
                                        na.rm = TRUE)),
                   CV = 100*abs(sd(.data$sim, na.rm = TRUE)/
                        mean(.data$sim, na.rm = TRUE)),
                   RMSE = mean(sqrt((.data$obs - .data$sim)^2),
                               na.rm = TRUE)) %>%
         ungroup() %>%
         as.data.frame()
  fit <- melt(fit, id.vars = c("lake", "metric", "variable", "nyear"))
  colnames(fit) <- c("lake", "metric", "variable", "nyear", "fit", "value")
  return(fit)
}

create_table <- function(df, 
                         metric_name, 
                         variable_breaks, 
                         variable_labels,
                         convert_units = "",
                         round_precision = NA,
                         lakes = c("Pleasant", "Long", "Plainfield")) {
  table_df <- df %>% 
              filter(.data$metric == metric_name) %>% 
              mutate(value = round(.data$value, digits = 2))
  if (convert_units == "meterTOft") {
    table_df$value <- NISTmeterTOft(table_df$value)
  }
  if (!is.na(round_precision)) {
    table_df$value <- round((1/round_precision)*table_df$value)/(1/round_precision)
  }
  if (variable_breaks == "sort") {
    variable_breaks <- sort(as.numeric(unique(table_df$variable)))
  }
  if (variable_labels == "breaks") {
    variable_labels <- variable_breaks
  } else if (variable_labels == "percent") {
    variable_labels <- sprintf("%d%%", variable_breaks)
  }
  table_df$variable  <- factor(table_df$variable,
                               levels = variable_breaks,
                               labels = variable_labels)
  table_df$fit <- NULL
  table_df <- dcast(table_df, lake~variable, value.var = "value")
  colnames(table_df) <- c("Lake", variable_labels)
  
  return(table_df)
}

```

# Overview
The goal of this exercise is to identify whether a 38 year timeseries is
sufficient to reasonably estimate the "true" hydrologic metrics. The "true"
values are of course unknowable, but we approximate them using metrics
calculated with the complete timeseries of data (1905-2018 for CSLS lakes;
1936-2001 for Devil's Lake). 

For each calculated metric, the **first plot** illustrates the 1) bias, 2)
precision, and 3) accuracy of hydrologic metrics calculated based on 100 random
samples of a continuous timeseries of lake levels at annual increments. The
error measurements displayed here are:

 1. **PBIAS**, or percent bias. Difference between simulated metric and "true"
 long-term metric relative to the long-term metric (units of percent). "Biased" 
 metrics in Kennard were those with mean bias > 30% (most were < 20%).
 2. **CV**, or coefficient of variation. Standard deviation of simulated metric
 relative to the mean of the simulated metric (units of percent). "Imprecise"
 metrics in Kennard were those with mean CV > 30% (most were < 25%).
 3. **RMSE**, or root mean squared error. Difference between simulated metric and
 "true" long-term metric as an absolute value (units are same as the metric, e.g.
 "feet" for median levels, "percent" for CV of levels, "months" for
 median duration).

The **second plot** contains a) box plots of the range of
values calculated in 100 random samples of 38 continuous years of lake levels,
b) a triangle representing the value calculated with the full timeseries, and c)
a square representing the value calculated using data from 1981-2018.

Lastly, **two tables** are displayed. The first shows the difference between
1981-2018 values of metrics vs. 1905-2018 values. The second shows the
difference between the mean value for a 38-year timeseries vs. 1905-2018 values.

<br>

```{r calc, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
# Metrics for all simulations with nyears = 38
short_fit <- monte_carlo_metrics %>% 
             filter(.data$nyear == 38) %>%
             group_by(.data$lake, .data$metric, .data$variable) %>%
             summarise(sim = median(.data$sim),
                       obs = mean(.data$obs)) %>%
             mutate(value = .data$sim - .data$obs) %>%
             ungroup() %>%
             as.data.frame() %>%
             select(.data$lake, .data$metric, .data$variable, .data$value)

df_box <- monte_carlo_metrics %>% 
          filter(.data$nyear == 38) %>%
          select(lake = .data$lake,
                 metric = .data$metric,
                 variable = .data$variable,
                 value = .data$sim,
                 nsim = .data$nsim)

# Metrics for 1905-2018
csls_metrics_obs <- monte_carlo_metrics %>%
                    select(.data$lake, .data$metric, .data$variable, .data$obs) %>%
                    unique()
df_point         <- csls_metrics_obs %>%
                    select(lake = .data$lake,
                           metric = .data$metric,
                           variable = .data$variable,
                           value = .data$obs)

# Metrics for all simulations
monte_carlo_metrics <- calculate_fit(monte_carlo_metrics)
monte_carlo_metrics <- monte_carlo_metrics %>%
                       mutate(value = ifelse(.data$metric %in% 
                                                c("median_level",
                                                  "exceedance_level",
                                                  "exceedance_range",
                                                  "median_rise_rate",
                                                  "median_fall_rate") &
                                                .data$fit == "RMSE",
                                              NISTmeterTOft(.data$value),
                                              .data$value))

# Metrics & fit for 1981-2018
csls_metrics <- calculate_metrics(csls_levels)
if (convert_to_depth) {
csls_metrics <- merge(csls_metrics, 
                             lake_bottom, 
                             by = "lake", 
                             all.x = TRUE)
csls_metrics <- csls_metrics %>%
                mutate(value = ifelse(.data$metric %in% c("median_level",
                                                          "exceedance_level"),
                                      .data$value - .data$bottom,
                                      .data$value))
csls_metrics$bottom <- NULL
}
df_point_short   <- csls_metrics
colnames(csls_metrics)[which(colnames(csls_metrics) == "value")] <- "sim"
csls_metrics     <- merge(csls_metrics, 
                          csls_metrics_obs,
                          by = c("lake", "metric", "variable"),
                          all.x = TRUE)
csls_fit         <- csls_metrics %>%
                    group_by(.data$lake, .data$metric, .data$variable) %>%
                    mutate(value = .data$sim - .data$obs) %>%
                    ungroup() %>%
                    as.data.frame() %>%
                    select(.data$lake, .data$metric, .data$variable, .data$value)
```

# Summary of Decisions

 1. **Magnitude:** Ok to use median levels and CV of levels.
 2. **Frequency:** Ok to use exceedance probability levels, but do not use ranges
 3. **Duration:** Must be calculated with entire timeseries
 4. **Rate of Change:** Ok to use all, but focus on monthly and seasonal rates
 (more robust than annual rates).

<br>

# Levels

For reference, here is the monthly record of lake levels at each lake. Note that
hydrologic metrics are calculated using the imputed levels for Pleasant, Long,
and Plainfield lake, but are calculated using the observed levels for Devils
Lake. The range of the y-axes are scaled such that 1 foot is the same on all
subplots.

```{r levels, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=5}
plot_levels(CSLSlevels::csls_levels, point_size = 1, convert_units = "meterTOft")
```

<br>

# Magnitude

## Median levels

**With a timeseries of 38 years,** bias is below 10%, variance is below 15%, and
accuracy is within 0.5 ft for all lakes.

**With longer timeseries,** variance and accuracy worsen for timeseries lengths
~45 years to 75 years at Long Lake, while other lakes see a steady, but not
dramatic, improvement. Bias is consistently low at all lakes.

**Decision:** it is acceptable to use a 38-year timeseries to calculate median
lake levels. All measures of error are within acceptable ranges and do not
sharply improve with longer timeseries.

**Implications of using 1981-2018 timeseries:** Median lake levels from
1981-2018 (1981-2001 for Devil’s Lake) are the highest of any 38-year window and
are also above the long-term median lake levels.

```{r median_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_level"
metric_title    <- ""
variable_title  <- "Month"
variable_breaks <- "sort"
variable_labels <- c("Annual","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec")
variable_colors <- c("#000000", brewer.pal(12, "Paired"))
value_title     <- "Median Level (ft)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 0.5))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r median_level_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title = "", variable_breaks, variable_labels, value_title,
           convert_units = "meterTOft", rotate_xticks = TRUE)
```

<br>

```{r median_level_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "1981-2018 Difference (ft)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (ft)")
```

## CV of levels

**With a timeseries of 38 years,** bias is below 20%, variance is below 30%, and
accuracy is within 0.03%. Devil’s Lake has noticeably higher bias and RMSE
compared to the Central Sands Lakes, while Pleasant Lake has noticeably higher
variance than other lakes.

**With longer timeseries,** in Pleasant Lake, bias increases and then decreases
again and variance markedly improves. All other metrics at all other lakes
either plateau or gradually decline.

**Decision:** it is acceptable to use a 38-year timeseries to calculate the
coefficient of variation in lake levels. All measures of error are within an
acceptable range and do not sharply improve with longer timeseries, except for
variance at Pleasant Lake.

**Implications of using 1981-2018 timeseries:** The CVs of lake levels are
nearly identical to the long-term CVs at Long and Devil’s Lakes but are among
the lowest observed CV’s at Pleasant and Plainfield Lakes.

```{r cv_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_level"
metric_title    <- ""
variable_title  <- "Month"
variable_breaks <- "sort"
variable_labels <- c("Annual","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec")
variable_colors <- c("#000000", brewer.pal(12, "Paired"))
value_title     <- "CV of Levels (%)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 0.03))
plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r cv_level_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title = "", variable_breaks, variable_labels, value_title,
           convert_units = "", ytick_precision = NA, same_range = FALSE, 
           rotate_xticks = TRUE)
```
<br>

```{r cv_level_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.01)
kableExtra::kable(table_df, caption = "1981-2018 Difference (%)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.01)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (%)")
```

<br>

# Frequency

## Exceedance probabilities

**With a timeseries of 38 years,** bias is below 5%, men variance is below 15%,
and accuracy is within 0.5 ft (Figure A3). Accuracy is just above 0.5 ft for the
10% exceedance probability at Devil’s Lake, but the 10% exceedance probability
level is not noticeably worse than other exceedance probabilities at other
lakes.

**With longer timeseries,** the bias for all lakes further improves. Variance at
Pleasant Lake and Devil’s Lake plateau with longer timeseries. Variance and
accuracy worsen before improving at Long Lake. All other metrics gradually but
steadily improve with longer timeseries.

**Decision:** it is acceptable to use a 38-year timeseries to calculate lake
levels associated with a 10%, 25%, 50%, 75% and 90% exceedance probability. All
measures of error are within the acceptable range and do not sharply improve
with longer timeseries. One exception to this might be 10% exceedance levels at
Devil’s Lake.

**Implications of using 1981-2018 timeseries:** The selection of 1981-2018
(1981-2001 for Devil’s Lake) causes some or all exceedance probability levels to
be higher than the long-term values. At Pleasant and Plainfield Lakes,
infrequent high levels (10% and 25% exceedance probabilities) are reasonably
close to their long-term values, but median levels and infrequent low levels
(75% and 90% exceedance probabilities) are high. At Long Lake, only the middle
exceedance levels (25%, 50%, 75%) are substantially higher than the long-term
values. At Devil’s Lake, all levels are higher than their long-term values.

```{r exceedance_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "exceedance_level"
metric_title    <- ""
variable_title  <- "Exceed. Prob"
variable_breaks <- "sort"
variable_labels <- "percent"
variable_colors <- c(brewer.pal(6, "Paired")[2:1],
                     "#000000",
                     brewer.pal(6, "Paired")[5:6])
value_title     <- "Lake Level (ft)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 0.5))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r exceedance_level_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title = "", variable_breaks, variable_labels, value_title,
           convert_units = "meterTOft")
```
<br>

```{r exceedance_level_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=5.5}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "1981-2018 Difference (ft)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (ft)")
```

## Exceedance probability ranges

**With a timeseries of 38 years,** bias is less than 25%, variance is within
40%, and accuracy is within 0.7 ft.

**With longer timeseries,** many measures of error noticeably improve, though
some (e.g., Pleasant Lake bias, Plainfield Lake variance) worsen before
improving and others plateau until much longer timeseries (e.g., accuracy at
Pleasant, Long, and Plainfield). Measures of error are generally worse for the
range between the 25% and 75% exceedance probabilities than they are for the
range between 10% and 90% exceedance probabilities, with the exception of
accuracy and Pleasant Lake (where both ranges are similar) and accuracy at
Plainfield and Devil’s Lakes (where 10%-90% are worse than 25%-75%).

**Decision:** it is not acceptable to use a 38-year timeseries to calculate the
range between the 25% and 75% exceedance probabilities. These ranges are not as
robust as the levels themselves and error is generally not within acceptable
ranges. However, it is acceptable to use the range between the 10% and 90%
exceedance probabilities, as measures of error for this range are more
consistently within acceptable levels.

**Implications of using 1981-2018 timeseries:** There is no systematic pattern
in whether exceedance probability ranges are above, below, or close to the
long-term value.

```{r exceedance_range_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "exceedance_range"
metric_title    <- ""
variable_title  <- "Exceed. Prob"
variable_breaks <- c("a_10_90","a_25_75","m_10_90","m_25_75")
variable_labels <- c("Annual 10%-90%", "Annual 25%-75%", 
                     "Monthly 10%-90%", "Monthly 25%-75%")
variable_colors <- c(brewer.pal(4, "Paired"))
value_title     <- "Lake Level Range (ft)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 0.5))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r exceedance_range_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title = "", variable_breaks, variable_labels, value_title,
           convert_units = "meterTOft", rotate_xticks = TRUE)
```
<br>

```{r exceedance_range_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=8, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "1981-2018 Difference (ft)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (ft)")
```

<br>

# Duration

## Median duration

**With a timeseries of 38 years,** bias is over 100%, variance is over 100% and
RMSE is as high as 20 months.

**With longer timeseries,** most measures of error improve, but none are within
acceptable ranges until very close to the full timeseries. Some measures of
error (e.g., Pleasant Lake 75% bias and accuracy, Devil’s Lake 10% variance)
remain high even at the longest period of record.

**Decision:** the full timeseries must be used to calculate median duration.
Some measures of error remain high even with timeseries as long as 90 years.

**Implications of using 1981-2018 timeseries:** The median durations calculated
using 1981-2018 timeseries (1981-2001 for Devil’s Lake) are reasonably close to
long-term values. However, per our evaluation of the time period of analysis,
there is too much uncertainty to reasonably use a short timeseries to estimate
typical duration at or in excess of infrequent high or low lake levels, so we
will not use these to describe the historical lake level regime.

```{r median_dur_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_dur"
metric_title    <- ""
variable_title  <- "Exceed. Prob"
variable_breaks <- "sort"
variable_labels <- "percent"
variable_colors <- c(brewer.pal(6, "Paired")[2:1],
                     brewer.pal(6, "Paired")[5:6])
value_title     <- "Months"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 1))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r median_dur_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title = "", variable_breaks, variable_labels, value_title,
           convert_units = "", same_range = FALSE)
```
<br>

```{r median_dur_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "1981-2018 Difference (months)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (months)")
```

## CV of duration

**With a timeseries of 38 years,** bias is within 50%, variance is within 50%,
and accuracy is within 125%.

**With longer timeseries,** most metrics improve, but none are within acceptable
ranges until very close to the full timeseries.

**Decision:** the full timeseries must be used to calculate the coefficient of
variation for duration. Some measures of error meet our targets at timeseries of
90 years, but the full period of record is necessary to meet the target for all
metrics on all lakes.

**Implications of using 1981-2018 timeseries:** The CV of durations calculated
using 1981-2018 timeseries (1981-2001 for Devil’s Lake) are reasonably close to
long-term values. However, per our evaluation of the time period of analysis,
there is too much uncertainty to reasonably use a short timeseries to estimate
typical duration at or in excess of infrequent high or low lake levels, so we
will not use these to describe the historical lake level regime.

```{r cv_dur_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_dur"
metric_title    <- ""
variable_title  <- "Exceed. Prob"
variable_breaks <- "sort"
variable_labels <- "percent"
variable_colors <- c(brewer.pal(6, "Paired")[2:1],
                     brewer.pal(6, "Paired")[5:6])
value_title     <- "CV of Duration (%)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 20))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r cv_dur_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title = "", variable_breaks, variable_labels, value_title,
           convert_units = "", same_range = FALSE)
```

```{r cv_dur_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)
kableExtra::kable(table_df, caption = "1981-2018 Difference (%)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (%)")
```

<br>

# Rate of change

## Median rise rate

**With a timeseries of 38 years,** bias is within 15%, variance is within 20%,
and accuracy is within 0.1 ft/time period.

**With longer timeseries,** bias does not substantially improve (except in
Devils Lake), but variance and accuracy gradually continue to decline.

**Decision:** it is acceptable to use a 38-year timeseries to calculate the
median rise/fall rate. All measures of error are within the acceptable range,
however monthly and seasonal rise/fall rates are more robust than annual
rise/fall rates.

**Implications of using 1981-2018 timeseries:** Rates of change appear to be
relatively insensitive to the selection of a wet period. Median rates of changes
are within 0.1 ft/time period of the long-term median rates of change.

```{r median_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_rise_rate"
metric_title    <- ""
variable_title  <- "Months"
variable_breaks <- "sort"
variable_labels <- "breaks"
variable_colors <- c(brewer.pal(6, "Paired")[c(1, 3, 5)])
value_title     <- "Rise Rate (ft/time period)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 0.1))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r median_rise_rate_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title, variable_breaks, variable_labels, value_title,
           convert_units = "meterTOft", same_range = FALSE)
```

```{r median_rise_rate_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "1981-2018 Difference (ft/time period)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (ft/time period)")
```

## CV of rise rate

**With a timeseries of 38 years,** bias is within 10%, variance is within 15%,
and accuracy is within 5% for monthly and seasonal rates. Accuracy of annual
rates is within 10%.

**With longer timeseries,** bias does not substantially improve (except in
Devils Lake), but variance and accuracy gradually continue to decline.

**Decision:** it is acceptable to use a 38-year timeseries to calculate the
coefficient of variation in rise/fall rate. All measures of error are within the
acceptable range, however monthly and seasonal rise/fall rates are more robust
than annual rise/fall rates.

**Implications of using 1981-2018 timeseries:** Rates of change appear to be
relatively insensitive to the selection of a wet period. The CV of rates of
change is sometimes similar and sometimes an outlier, but there is no clear
pattern to these differences.

```{r cv_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_rise_rate"
metric_title    <- ""
variable_title  <- "Months"
variable_breaks <- "sort"
variable_labels <- "breaks"
variable_colors <- c(brewer.pal(6, "Paired")[c(1, 3, 5)])
value_title     <- "CV of Rise Rate (%)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 5))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r cv_rise_rate_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title, variable_breaks, variable_labels, value_title,
           convert_units = "", same_range = FALSE)
```

```{r cv_rise_rate_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "1981-2018 Difference (%)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (%)")
```

## Median fall rate

Same observations as median rise rate.

```{r median_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_fall_rate"
metric_title    <- ""
variable_title  <- "Months"
variable_breaks <- "sort"
variable_labels <- "breaks"
variable_colors <- c(brewer.pal(6, "Paired")[c(1, 3, 5)])
value_title     <- "Fall Rate (ft/time period)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 0.1))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r median_fall_rate_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title, variable_breaks, variable_labels, value_title,
           convert_units = "meterTOft", same_range = FALSE)
```

```{r median_fall_rate_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.01)
kableExtra::kable(table_df, caption = "1981-2018 Difference (ft/time period)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.01)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (ft/time period)")
```

## CV of fall rate

Same observations as CV of rise rate.

```{r cv_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_fall_rate"
metric_title    <- ""
variable_title  <- "Months"
variable_breaks <- "sort"
variable_labels <- "breaks"
variable_colors <- c(brewer.pal(6, "Paired")[c(1, 3, 5)])
value_title     <- "CV of Fall Rate (%)"
hlines          <- data.frame(fit = c("PBIAS", "CV", "RMSE"),
                              value = c(30, 30, 5))

plot_sensitivity(monte_carlo_metrics, metric_name, metric_title, variable_title, 
                 variable_breaks, variable_labels, variable_colors, text_size,
                 hlines = hlines)
```

<br>

```{r cv_fall_rate_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
plot_nyear(df_box, df_point, df_point_short, metric_name, metric_title = "",
           variable_title, variable_breaks, variable_labels, value_title,
           convert_units = "", same_range = FALSE)
```

```{r cv_fall_rate_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
table_df <- create_table(csls_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "1981-2018 Difference (%)")

table_df <- create_table(short_fit, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)
kableExtra::kable(table_df, caption = "Mean 38 year Difference (%)")
```
