---
title: "A Look Back at Long Lake"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{A Look Back at Long Lake}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSlevels)
library(dplyr)
library(lubridate)
library(magick)
library(cowplot)
library(ggplot2)
library(ggspatial)

# Parameters
text_size   <- 10

# Photo observations (location & dates)
obs_loc      <- data.frame(x = -89.448499, y = 44.204690)
photo_dates1 <- as_datetime(ymd(c("2006-07-01", "2006-08-01", "2006-09-01")))
photo_dates2 <- as_datetime(ymd(c("2009-08-01", "2009-09-01", "2009-10-01")))
photo_dates3 <- as_datetime(ymd("1964-10-01"))

# Lake level data
lake        <- "Long"
csls_levels <- CSLSlevels::csls_levels
lake_levels <- subset(csls_levels, lake_name == sprintf("%s Lake", lake))

lake_levels_2000s  <- lake_levels %>%
                      filter(year(obs_mo) > 2003,
                             year(obs_mo) < 2012)
lake_levels_1960s  <- lake_levels %>%
                      filter(year(obs_mo) > 1957,
                             year(obs_mo) < 1966)
photo_levels_2000s <- lake_levels %>%
                      filter(as_datetime(obs_mo) %in% 
                               c(photo_dates1, photo_dates2),
                             !is.na(lake_lev_obs))
photo_levels_1960s <- lake_levels %>%
                      filter(as_datetime(obs_mo) %in% 
                               c(photo_dates1, photo_dates2),
                             !is.na(lake_lev_obs))

# Exceedance probabilities
ranked_annual      <- rank_lake_levels(lake_levels, 
                                       window_start = 1981, 
                                       window_end = 2018,
                                       lake_names = "Long Lake")
photo_levels_1960s$prob <- approx(x = ranked_annual$level,
                                  y = ranked_annual$prob,
                                  xout = photo_levels_1960s$lake_lev_obs)$y
```

## FAQ: How could lake levels be lower at Long Lake in 1964 than in 2007?

One aspect of Long Lake's history that can be puzzling to those who remember how
low lake levels were in the early 2000s is that levels are recorded as being
even lower in 1964. Personal photos from homeowners taken from the eastern
shoreline depict the Long Lake as a dense meadow of vegetation with no visible
standing water in 2009. The year 2007 saw lower lake levels and is particularly
memorable as a year that the lake "dried up" as seen from the shoreline. By
contrast, a personal photo from a homewoner dated 1964 shows standing water
clearly visible from the northern shoreline. Yet 1964 lake levels were
documented as being 11 cm lower than in 2007, 28 cm lower than in 2006, and 40
cm lower than 2009.

<center>
**How can this be?**
</center>

<br>

To investigate this, we paired historic lake level observations with detailed
lake bathymetry data collected as part of the WI DNR's Central Sands Lakes Study
to generate maps of lake depth. We then paired these lake depth maps with the
views from the shoreline that homeowners captured in their personal photos.

## What do we know about lake levels during these years?

Let's start by taking a step back to look at what we know about lake level
behavior leading up to each of these dates. We have much less data leading up to
1964 than we do in the early 2000s, but the data we do have shows that lake
levels were in a steep decline for several years leading up to 1964.



```{r levels, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 3}
plot_levels <- ggplot() +
               geom_line(data = lake_levels_2000s,
                         aes(x = as.Date(obs_mo), y = lake_lev_pred)) +
               geom_point(data = lake_levels_2000s,
                          aes(x = as.Date(obs_mo), y = lake_lev_obs)) +
               geom_point(data = photo_levels_1960s,
                          aes(x = as.Date(obs_mo), y = lake_lev_obs),
                          color = "red") +
               theme_bw() +
               labs(y = "Lake Level (mamsl)",
                    x = "") +
               theme(text = element_text(family = "Segoe UI Semilight",
                                         size = text_size),
               panel.grid.major = element_blank())

plot_exceed <- ggplot() +
               geom_line(data = ranked_annual,
                         aes(x = prob, y = level),
                         color = "black") +
               geom_point(data = photo_levels_1960s,
                          aes(x = prob, y = lake_lev_obs),
                          color = "red",
                          size = 2) +
               scale_x_continuous(expand = c(0, 0)) +
               labs(x = "Exceedance Probability (%)",
                    y = "Lake Level (mamsl)") +
               theme_bw() +
               theme(text = element_text(family = "Segoe UI Semilight",
                                        size = text_size))

plot_grid(plot_levels, plot_exceed)
```


Some of the lowest lake levels at Long Lake ever recorded occurred between 2006
and 2010. Brian Wolf, a homeowoner on Long Lake, collected photodocumentation of
lake levels during this time. According to George Kraft (with input from Brian
Wolf):

> *Brian and family bought their place on east end of Long Lake in April 2005.
> They were told the lake is a little low but it always comes back. That’s where
> the photodocumentation begins.*

Here, we're comparing all photos with a) a known location and visible shoreline,
b) a known month and year, and c) an observation within 2 months of that date to
d) bathymetry maps of the lake at the observed level.

<br>

There are two periods of time that meet this criteria: Summer 2006 and late
Summer/Fall 2009 (the red points). Lake levels at these two times represent some
of the lowest levels captured in the lake level exceedance curve for Long Lake
(>95% exceedance probability). Lake levels were rapidly falling in Summer 2006
and were beginning to rise in late Summer/Fall 2009, but the the recorded lake
levels are approximately the same at both dates (within 12 cm of each other).
This provides an interesting comparison of how vegetation at Long Lake responded
to the same water level when levels were falling vs. rising.


## Summer 2006

These photos were taken by homeowner Brian Wolf on July 29, 2006 from the east
side of Long Lake looking west. At this time, lake levels had been rapidly
falling for at least a year and lily pads are visible in some regions. The areas
with lily pads approximately correspond with the regions where depth is over 0.5
m.

```{r pic01, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 3}
im01         <- ggdraw() + draw_image("long_2006_07_im1.jpg", scale = 0.9)
im02         <- ggdraw() + draw_image("long_2006_07_im2.jpg", scale = 0.9)

plot_grid(im01, im02)
```

```{r bathy01, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 4}
lake_obs    <- lake_levels %>%
               filter(as_datetime(obs_mo) %in% photo_dates1)
lake_level  <- mean(lake_obs$lake_lev_obs, na.rm = TRUE)
bathy1       <- plot_lake_depths(lake, 
                                lake_level,
                                relative_depth_fill = TRUE,
                                title_name = sprintf("%s Lake", lake),
                                legend_position = "right",
                                text_size = text_size)+
                geom_spatial_point(data = obs_loc, 
                                   aes(x, y),
                                   crs = 4326,
                                   color = "red",
                                   shape = 8,
                                   size = 2.5)
bathy2       <- plot_lake_depths(lake, 
                                lake_level,
                                relative_depth_fill = FALSE,
                                title_name = sprintf("%s Lake", lake),
                                legend_position = "right",
                                text_size = text_size) +
                geom_spatial_point(data = obs_loc, 
                                   aes(x, y),
                                   crs = 4326,
                                   color = "red",
                                   shape = 8,
                                   size = 2.5)
bathy2
bathy1
```

## Summer 2009

George Kraft also shared photos from homeowner Brian Wolf taken on September 27,
2009 from the east side of Long Lake looking west. Lake levels at this time are
on their way up, but approximately the same as in Summer 2006 (only 12 cm
higher). Open water is visible approximately where the bathymetry maps indicate
water depth is over 0.5 m, but the surrounding area is filled with dense,
mid-height (~3-5 ft tall) vegetation.

```{r pic02, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 3}
im01        <- ggdraw() + draw_image("long_2009_09.jpg", scale = 0.9)
im02        <- ggdraw() + draw_image("long_2009_10.jpg", scale = 0.9)

plot_grid(im01, im02)
```

```{r bathy02, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 4}
lake_obs    <- lake_levels %>%
               filter(as_datetime(obs_mo) %in% photo_dates2)
lake_level  <- mean(lake_obs$lake_lev_obs, na.rm = TRUE)
bathy1       <- plot_lake_depths(lake, 
                                lake_level,
                                relative_depth_fill = TRUE,
                                title_name = sprintf("%s Lake", lake),
                                legend_position = "right",
                                text_size = text_size)+
                geom_spatial_point(data = obs_loc, 
                                   aes(x, y),
                                   crs = 4326,
                                   color = "red",
                                   shape = 8,
                                   size = 2.5)
bathy2       <- plot_lake_depths(lake, 
                                lake_level,
                                relative_depth_fill = FALSE,
                                title_name = sprintf("%s Lake", lake),
                                legend_position = "right",
                                text_size = text_size) +
                geom_spatial_point(data = obs_loc, 
                                   aes(x, y),
                                   crs = 4326,
                                   color = "red",
                                   shape = 8,
                                   size = 2.5)
bathy2
bathy1
```


## References
