---
title: "CSLS Lake Metrics"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Lake Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSlevels)
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
library(extrafont)
library(lubridate)
library(NISTunits)
library(stringr)
library(reshape2)

# Parameters
text_size         <- 12
lakes             <- c("Pleasant", "Long", "Plainfield")

# Data
hist_levels       <- CSLSlevels::hist_levels
hist_levels_short <- hist_levels %>%
                     filter(year(.data$date) >= 1981,
                            year(.data$date) <= 2018)
hist_metrics       <- calculate_metrics(hist_levels_short)
hist_metrics_short <- CSLSlevels::hist_metrics_short %>%
                      filter(.data$lake != "Devils",
                             .data$variable != "b50") %>%
                      mutate(variable = ifelse(as.character(.data$variable) ==
                                                 "a50", "50", .data$variable)) %>%
                      group_by(.data$lake, .data$metric, .data$variable) %>%
                      summarise(median = median(.data$value),
                                sd = sd(.data$value)) %>%
                      ungroup()
hist_metrics_long  <- CSLSlevels::hist_metrics_long %>%
                      filter(.data$lake != "Devils") %>%
                      group_by(.data$lake, .data$metric, .data$variable) %>%
                      summarise(median = median(.data$value),
                                sd = sd(.data$value)) %>%
                      ungroup()

# Functions
create_table <- function(df, 
                         metric_name, 
                         variable_breaks, 
                         variable_labels,
                         convert_units = "",
                         round_precision = NA) {
  table_df <- df %>% 
              filter(.data$metric == metric_name) %>%
              select(.data$lake, .data$variable, .data$median, .data$sd)
  
  if (convert_units == "meterTOft") {
    table_df <- table_df %>% mutate_at(c("median", "sd"), NISTmeterTOft)
  }
  if (!is.na(round_precision)) {
    table_df$median <- round((1/round_precision)*table_df$median)/(1/round_precision)
    table_df$sd     <- round((1/round_precision)*table_df$sd)/(1/round_precision)
  }
  if (variable_breaks == "sort") {
    variable_breaks <- sort(as.numeric(unique(table_df$variable)))
  }
  if (variable_labels == "breaks") {
    variable_labels <- variable_breaks
  } else if (variable_labels == "percent") {
    variable_labels <- sprintf("%d%%", variable_breaks)
  }
  table_df$variable  <- factor(table_df$variable,
                               levels = variable_breaks,
                               labels = variable_labels)
  
  table_df$text <- sprintf("%s \U00B1 %s", table_df$median, table_df$sd)

  table_df <- dcast(table_df, lake~variable, value.var = "text")
  colnames(table_df) <- c("Lake", variable_labels)
  
  return(table_df)
}
```

<br>

# Historical Lake Levels

Estimated lake levels from 1938-2019 for each study lake (black line) with the
mean value of all estimations +/- one standard deviation shown in dark grey
(captures 68.2% of estimated values), the mean value of all estimations +/- two
standard deviations shown in light grey (captures 95% of estimated values), and
observations shown as points. Devils Lake (1936-2001) is also shown for
comparison.

```{r levels01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=5}
plot_levels(hist_levels, 
            legend_pos = "top", 
            show_obs_type = TRUE, 
            color_vals = c("black", "#1B9E77", "#7570B3", "#D95F02"), 
            line_size = 0.7, 
            point_size = 1.5, 
            npretty_breaks = 5,  
            convert_units = "meterTOft", 
            grid_off = FALSE, 
            show_uncertainty = TRUE,
            text_size = text_size,
            lakes = c("Pleasant", "Long", "Plainfield"),
            force_range = FALSE)
```

<br>

# Time Period of Analysis 

Estimated lake levels for each lake (black line) with observations (red points)
from 1981-2018 (Devils Lake: 1981-2001). Light grey shaded area represents the
range between the infrequent low lake level (90% exceedance probability) and
infrequent high lake level (10% exceedance probability). Solid grey line
represents the median (50% exceedance probability).

```{r levels02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=5}
plot_levels(hist_levels_short, 
            grid_off = FALSE, 
            point_size = 1.5, 
            line_size = 0.7,
            legend_pos = 'top', 
            show_obs_type = FALSE,
            convert_units = "meterTOft",
            all_probs = TRUE,
            npretty_breaks = 5,
            color_vals = c("black", "#c00000"),
            hline_color = c("grey70"),
            text_size = text_size,
            force_range = FALSE)
```

<br>

# Historical Lake Level Regimes

## 1. Magnitude
Range in lake elevations (ft) for all monthly lake elevations (1981-2018,
“Overall”) as well as for each month from 1981-2018 for each study lake. All
y-axes cover the same range so differences in lake elevation are comparable
across plots. On this box and whisker plot, the line in the middle of each box
represents the median value. The lower and upper sides of each box represent the
first and third quartiles (25th and 75th percentiles), respectively. The upper
whisker extends from the third quartile to the largest value no more than 1.5x
the interquartile range (difference between the first and third quartile), while
the lower whisker extends from the first quartile to the smallest value no
smaller than 1.5x the interquartile range. Points represent outliers which fall
beyond the maximum extent of the whiskers.

```{r magnitude_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=5, fig.height=5,}
plot_magnitude(hist_levels_short, text_size = text_size, force_range = FALSE)
```

<br>

```{r median_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_level"
metric_title    <- "**Median Lake Level (ft)**"
variable_breaks <- "sort"
variable_labels <- c("Annual","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r cv_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_level"
metric_title    <- "**CV of Lake Levels (%)**"
variable_breaks <- "sort"
variable_labels <- c("Annual","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r frequency, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=4}
plot_frequency(hist_levels_short, probs = c(10, 50, 90), text_size = text_size) + theme(panel.spacing.x = unit(6, "mm"))
```

<br>

```{r exceedance_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6.5}
metric_name     <- "exceedance_level"
metric_title    <- "**Exceedance Probability Lake Levels (ft)**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

```{r exceedance_level_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6.5}
elev_area_vol <- CSLSdata::elev_area_vol
lake_bottom <- data.frame(lake = lakes,
                          bottom = c(291.1426, 332.8622, 332.0755))

table_df      <- hist_metrics_short %>%
                 filter(.data$metric == metric_name) %>%
                 mutate(area_m2 = NA,
                       area_high = NA,
                       area_low = NA,
                       vol_m3 = NA,
                       vol_high = NA,
                       vol_low = NA) %>%
                 left_join(lake_bottom, by = "lake")
table_df$metric <- NULL

# Translate elevations to area, vol, and depths
for (lake in unique(elev_area_vol$lake)) {
    this_elev_area_vol <- elev_area_vol %>% filter(.data$lake == !!lake)
    f_elev_area        <- approxfun(x = this_elev_area_vol$elev_m,
                                    y = this_elev_area_vol$area_m2)
    f_elev_vol         <- approxfun(x = this_elev_area_vol$elev_m,
                                    y = this_elev_area_vol$vol_m3)
    table_df           <- table_df %>%
                          mutate(area_m2 = ifelse(.data$lake == !!lake,
                                                  f_elev_area(.data$median),
                                                  .data$area_m2),
                                 area_high = ifelse(.data$lake == !!lake,
                                                  f_elev_area(.data$median + .data$sd),
                                                  .data$area_high),
                                 area_low = ifelse(.data$lake == !!lake,
                                                  f_elev_area(.data$median - .data$sd),
                                                  .data$area_low),
                                 vol_m3 = ifelse(.data$lake == !!lake,
                                                  f_elev_vol(.data$median),
                                                  .data$vol_m3),
                                 vol_high = ifelse(.data$lake == !!lake,
                                                  f_elev_vol(.data$median + .data$sd),
                                                  .data$vol_high),
                                 vol_low = ifelse(.data$lake == !!lake,
                                                  f_elev_vol(.data$median - .data$sd),
                                                  .data$vol_low),
                                 mean_depth = .data$vol_m3/.data$area_m2,
                                 mean_high = .data$vol_high/.data$area_high,
                                 mean_low = .data$vol_low/.data$area_low,
                                 max_depth = .data$median - .data$bottom,
                                 max_high = .data$median + .data$sd - .data$bottom,
                                 max_low = .data$median - .data$sd - .data$bottom)
}

table_df$area_diff <- apply(data.frame(high = table_df$area_high - table_df$area_m2,
                                       low = table_df$area_m2 - table_df$area_low),
                            1, max)
table_df$vol_diff  <- apply(data.frame(high = table_df$vol_high - table_df$vol_m3,
                                       low = table_df$vol_m3 - table_df$vol_low),
                            1, max)
table_df$mean_diff <- apply(data.frame(high = table_df$mean_high - table_df$mean_depth,
                                       low = table_df$mean_depth - table_df$mean_low),
                            1, max)
table_df$max_diff <- apply(data.frame(high = table_df$max_high - table_df$max_depth,
                                       low = table_df$max_depth - table_df$max_low),
                            1, max)
table_df <- table_df %>%
            select(lake = .data$lake,
                   variable = .data$variable,
                   elev = .data$median,
                   elev_diff = .data$sd,
                   area = .data$area_m2,
                   area_diff = .data$area_diff,
                   vol = .data$vol_m3,
                   vol_diff = .data$vol_diff,
                   mean_depth = .data$mean_depth,
                   mean_diff = .data$mean_diff,
                   max_depth = .data$max_depth,
                   max_diff = .data$max_diff)

# Convert to English units
table_df <- table_df %>%
            mutate_at(c("elev", "elev_diff", "mean_depth", "mean_diff", "max_depth", "max_diff"),
                      NISTmeterTOft) %>%
            mutate_at(c("area", "area_diff"), NISTsqrMeterTOacre) %>%
            mutate_at(c("vol", "vol_diff"), NISTcubMeterTOacreFt)

# Round appropriately
round_precision <- 0.1
for (col in c("elev", "elev_diff", "mean_depth", "mean_diff", "max_depth", 
              "max_diff", "area", "area_diff")) {
  table_df[, col] <- round((1/round_precision)*table_df[, col])/(1/round_precision)

}
round_precision <- 1
for (col in c("vol", "vol_diff")) {
  table_df[, col] <- round((1/round_precision)*table_df[, col])/(1/round_precision)

}

# Convert to string
table_df <- table_df %>%
            mutate(elev = sprintf("%s \U00B1 %s", .data$elev, .data$elev_diff),
                   area = sprintf("%s \U00B1 %s", .data$area, .data$area_diff),
                   vol = sprintf("%s \U00B1 %s", .data$vol, .data$vol_diff),
                   mean_depth = sprintf("%s \U00B1 %s", .data$mean_depth, .data$mean_diff),
                   max_depth = sprintf("%s \U00B1 %s", .data$max_depth, .data$max_diff)) %>%
            select(.data$lake, .data$variable, .data$elev, .data$area, 
                   .data$vol, .data$mean_depth, .data$max_depth)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

## 2. Frequency

```{r num_exceedances_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "num_dur_decade"
metric_title    <- "**Exceedances per Decade**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

```{r depart_median_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "depart_median"
metric_title    <- "**Probability within 1 ft of median (%)**"
variable_breaks <- c("0.3m", "-0.3m")
variable_labels <- c("+1ft", "-1ft")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

## 3. Duration

```{r duration, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=7}
plot_duration(hist_levels_short, 
              probs = c(10, 25, 50, 75, 90), 
              show_median = TRUE, 
              bin_size = 1, 
              text_size = text_size,
              max_months = 12)
```

<br>

```{r median_dur_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_dur"
metric_title    <- "**Median Duration (months)**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r cv_dur_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_dur"
metric_title    <- "**CV of Duration (%)**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

## 4. Rate of change

```{r rate_of_change, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=7}
plot_rate_change(hist_levels_short, text_size = text_size)
```

<br>

```{r median_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_rise_rate"
metric_title    <- "**Median Rise Rate (ft/time period)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r cv_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_rise_rate"
metric_title    <- "**CV of Rise Rate (%)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r median_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_fall_rate"
metric_title    <- "**Median Fall Rate (ft/time period)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r cv_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_fall_rate"
metric_title    <- "**CV of Fall Rate (%)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

## 5. Timing

```{r timing, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4, fig.height=3}
plot_df <- hist_metrics_short %>%
           filter(.data$metric == "exceedance_range") %>%
           mutate(time = str_sub(.data$variable, 1, 1),
                  range = str_sub(.data$variable, 3, -1),
                  median = NISTunits::NISTmeterTOft(.data$median),
                  sd = NISTunits::NISTmeterTOft(.data$sd)) %>%
           mutate(range = str_replace(.data$range, "10_90", "10%-90%")) %>%
           mutate(range = str_replace(.data$range, "25_75", "25%-75%")) %>%
           mutate(text = sprintf("%s \U00B1 %s", .data$median, .data$sd)) %>%
           # filter(.data$range != "25%-75%") %>%
           select(.data$lake, .data$time, .data$range, .data$median, .data$sd)
plot_df <- dcast(plot_df, lake+range~time, value.var = "median")


ggplot(plot_df) +
  geom_point(aes(x = a, y = m, shape = lake, color = range), 
             # color = "black",
             size = 2) +
  geom_abline(slope = 1, intercept = 0) +
  labs(x = "Annual Range (ft)",
       y = "Monthly Range (ft)",
       shape = "") +
  theme_bw() +
  scale_color_manual(name = "",
                     breaks = c("10%-90%", "25%-75%"),
                     values = c("black", "grey70")) +
  theme(text = element_text(family = "Segoe UI Semilight",
                            size = text_size),
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        legend.box = "vertical",
        legend.margin = margin(0,0,0,0))
```

<br>

```{r exceedance_range_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "exceedance_range"
metric_title    <- "**Exceedance Probability Range (ft)**"
variable_breaks <- c("a_10_90","a_25_75","m_10_90","m_25_75")
variable_labels <- c("Annual 10%-90%", "Annual 25%-75%", 
                     "Monthly 10%-90%", "Monthly 25%-75%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
