---
title: "CSLS Historical Hydrologic Metrics (1981-2018)"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{CSLS Historical Hydrologic Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
# Overview

This file presents the hydrologic metrics from 1981-2018 based on all 4000 monte carlo predictions of historical lake level time series.

<br>

```{r setup, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(CSLSlevels)
library(dplyr)
library(lubridate)
library(reshape2)
library(ggplot2)
library(extrafont)
library(NISTunits)
library(stringr)

# PARAMETERS
text_size         <- 12
lakes             <- c("Pleasant", "Long", "Plainfield")

# DATA
# Lake level timeseries: Monthly 1981-2018
hist_levels_short  <- CSLSlevels::hist_levels %>%
                      filter(year(.data$date) >= 1981,
                             year(.data$date) <= 2018)

# Lake level timeseries: Seasonal (3-mo) 1981-2018
hist_levels_seas   <- hist_levels_short %>%
                      filter(.data$lake != "Devils") %>%
                      mutate(month = month(.data$date)) %>%
                      group_by(.data$lake) %>%
                      arrange(.data$date) %>%
                      mutate(level = as.numeric(stats::filter(.data$level,
                                                              rep(1/3, 3),
                                                              sides = 2))) %>%
                      ungroup() %>%
                      filter(.data$month %in% c(1,4,7,10))

# Lake level timeseries: Seasonal (3-mo) 1981-2018
hist_metrics_short <- CSLSlevels::hist_metrics_short %>%
                      filter(.data$lake != "Devils") %>%
                      group_by(.data$lake, .data$metric, .data$variable) %>%
                      summarise(median = median(.data$value),
                                sd = sd(.data$value)) %>%
                      ungroup()

# Convert number of exceedances to recurrence interval
recurr_int         <- hist_metrics_short %>%
                      filter(.data$metric %in% c("num_2yr_decade", 
                                                 "num_2yr_decade_seas")) %>%
                      mutate(metric = ifelse(.data$metric == "num_2yr_decade",
                                             "recurr_int_2yr", 
                                             "recurr_int_2yr_seas"),
                             median = 10/.data$median,
                             sd = ifelse(is.infinite(10/.data$sd),
                                         NA, 10/.data$sd))
hist_metrics_short <- rbind(hist_metrics_short, recurr_int)
```

```{r table_fcn, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
create_table <- function(df, metric_name, variable_breaks, variable_labels,
                         convert_units = "", round_precision = NA) {
  table_df <- df %>% 
              filter(.data$metric == metric_name) %>%
              select(.data$lake, .data$variable, .data$median, .data$sd)
  
  if (convert_units == "meterTOft") {
    table_df <- table_df %>% mutate_at(c("median", "sd"), NISTmeterTOft)
  }
  if (!is.na(round_precision)) {
    table_df$median <- round((1/round_precision)*table_df$median)/
                       (1/round_precision)
    table_df$sd     <- round((1/round_precision)*table_df$sd)/
                       (1/round_precision)
  }
  if (variable_breaks == "sort") {
    variable_breaks <- sort(as.numeric(unique(table_df$variable)))
  }
  if (variable_labels == "breaks") {
    variable_labels <- variable_breaks
  } else if (variable_labels == "percent") {
    variable_labels <- sprintf("%d%%", variable_breaks)
  }
  table_df$variable  <- factor(table_df$variable,
                               levels = variable_breaks,
                               labels = variable_labels)
  
  table_df$text <- sprintf("%s \U00B1 %s", table_df$median, table_df$sd)

  table_df <- dcast(table_df, lake~variable, value.var = "text")
  colnames(table_df) <- c("Lake", variable_labels)
  
  return(table_df)
}
```

# Historical Lake Levels

Estimated lake levels from 1938-2019 for each study lake (black line) with the
mean value of all estimations +/- one standard deviation shown in dark grey
(captures 68.2% of estimated values), the mean value of all estimations +/- two
standard deviations shown in light grey (captures 95% of estimated values), and
observations shown as points.

```{r levels01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=5}
plot_levels(CSLSlevels::hist_levels, 
            legend_pos = "top", 
            show_obs_type = TRUE, 
            color_vals = c("black", "#1B9E77", "#7570B3", "#D95F02"), 
            line_size = 0.7, 
            point_size = 1.5, 
            npretty_breaks = 5,  
            convert_units = "meterTOft", 
            grid_off = FALSE, 
            show_uncertainty = TRUE,
            text_size = text_size,
            lakes = lakes,
            force_range = FALSE)
```

<br>

# Time Period of Analysis 

Estimated lake levels for each lake (black line) with observations (red points)
from 1981-2018. Light grey shaded area represents the range between the
infrequent low lake level (90% exceedance probability) and infrequent high lake
level (10% exceedance probability). Solid grey line represents the median (50%
exceedance probability).

```{r levels02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=5}
plot_levels(hist_levels_short, 
            grid_off = FALSE, 
            point_size = 1.5, 
            line_size = 0.7,
            legend_pos = 'top', 
            show_obs_type = FALSE,
            convert_units = "meterTOft",
            all_probs = TRUE,
            npretty_breaks = 5,
            color_vals = c("black", "black"),
            hline_color = c("grey70"),
            text_size = text_size,
            force_range = FALSE) +
  geom_line(data = hist_levels_seas, 
            aes(x = .data$date, 
                y = NISTunits::NISTmeterTOft(.data$level)),
            color = "#c00000")
```

<br>

# Historical Lake Level Regimes
## 1. Magnitude
### Median Levels
```{r magnitude_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=5, fig.height=5}
plot_magnitude(hist_levels_short, text_size = text_size, force_range = FALSE)
```

<br>

```{r median_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_level"
metric_title    <- "**Median Lake Level (ft)**"
variable_breaks <- "sort"
variable_labels <- c("Annual","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### CV of Max Lake Depth

```{r cv_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_level"
metric_title    <- "**CV of Lake Levels (%)**"
variable_breaks <- "sort"
variable_labels <- c("Annual","Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul",
                    "Aug", "Sep", "Oct", "Nov", "Dec")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Exceedance Probabilities

```{r frequency_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=4}
plot_frequency(hist_levels_short, probs = c(10, 50, 90), text_size = text_size) + theme(panel.spacing.x = unit(6, "mm"))
```
<br>

Example of highlighting exceedances of infrequent high:
```{r frequency_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=4}
plot_levels(hist_levels_short, 
            grid_off = FALSE, 
            point_size = 1.5, 
            line_size = 0.7,
            legend_pos = 'top', 
            show_obs_type = FALSE,
            convert_units = "meterTOft",
            all_probs = FALSE,
            npretty_breaks = 5,
            color_vals = c("black", "black"),
            hline_color = "#7030A0",
            text_size = text_size,
            force_range = FALSE,
            show_exceeds = 10)
```
<br>

Example of highlighting exceedances of infrequent low:

```{r frequency_03, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=4}
plot_levels(hist_levels_short, 
            grid_off = FALSE, 
            point_size = 1.5, 
            line_size = 0.7,
            legend_pos = 'top', 
            show_obs_type = FALSE,
            convert_units = "meterTOft",
            all_probs = FALSE,
            npretty_breaks = 5,
            color_vals = c("black", "black"),
            hline_color = "#7030A0",
            text_size = text_size,
            force_range = FALSE,
            show_exceeds = 90)
```

<br>

```{r exceedance_level_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6.5}
metric_name     <- "exceedance_level"
metric_title    <- "**Exceedance Probability Lake Levels (ft)**"
variable_breaks <- "sort"
variable_labels <- "percent"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

```{r exceedance_level_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6.5}
elev_area_vol <- CSLSdata::elev_area_vol
lake_bottom <- data.frame(lake = lakes,
                          bottom = c(291.1426, 332.8622, 332.0755))

table_df      <- hist_metrics_short %>%
                 filter(.data$metric == metric_name) %>%
                 mutate(area_m2 = NA,
                       area_high = NA,
                       area_low = NA,
                       vol_m3 = NA,
                       vol_high = NA,
                       vol_low = NA) %>%
                 left_join(lake_bottom, by = "lake")
table_df$metric <- NULL

# Translate elevations to area, vol, and depths
for (lake in unique(elev_area_vol$lake)) {
    this_elev_area_vol <- elev_area_vol %>% filter(.data$lake == !!lake)
    f_elev_area        <- approxfun(x = this_elev_area_vol$elev_m,
                                    y = this_elev_area_vol$area_m2)
    f_elev_vol         <- approxfun(x = this_elev_area_vol$elev_m,
                                    y = this_elev_area_vol$vol_m3)
    table_df           <- table_df %>%
                          mutate(area_m2 = ifelse(.data$lake == !!lake,
                                                  f_elev_area(.data$median),
                                                  .data$area_m2),
                                 area_high = ifelse(.data$lake == !!lake,
                                                  f_elev_area(.data$median + .data$sd),
                                                  .data$area_high),
                                 area_low = ifelse(.data$lake == !!lake,
                                                  f_elev_area(.data$median - .data$sd),
                                                  .data$area_low),
                                 vol_m3 = ifelse(.data$lake == !!lake,
                                                  f_elev_vol(.data$median),
                                                  .data$vol_m3),
                                 vol_high = ifelse(.data$lake == !!lake,
                                                  f_elev_vol(.data$median + .data$sd),
                                                  .data$vol_high),
                                 vol_low = ifelse(.data$lake == !!lake,
                                                  f_elev_vol(.data$median - .data$sd),
                                                  .data$vol_low),
                                 mean_depth = .data$vol_m3/.data$area_m2,
                                 mean_high = .data$vol_high/.data$area_high,
                                 mean_low = .data$vol_low/.data$area_low,
                                 max_depth = .data$median - .data$bottom,
                                 max_high = .data$median + .data$sd - .data$bottom,
                                 max_low = .data$median - .data$sd - .data$bottom)
}

table_df$area_diff <- apply(data.frame(high = table_df$area_high - table_df$area_m2,
                                       low = table_df$area_m2 - table_df$area_low),
                            1, max)
table_df$vol_diff  <- apply(data.frame(high = table_df$vol_high - table_df$vol_m3,
                                       low = table_df$vol_m3 - table_df$vol_low),
                            1, max)
table_df$mean_diff <- apply(data.frame(high = table_df$mean_high - table_df$mean_depth,
                                       low = table_df$mean_depth - table_df$mean_low),
                            1, max)
table_df$max_diff <- apply(data.frame(high = table_df$max_high - table_df$max_depth,
                                       low = table_df$max_depth - table_df$max_low),
                            1, max)
table_df <- table_df %>%
            select(lake = .data$lake,
                   variable = .data$variable,
                   elev = .data$median,
                   elev_diff = .data$sd,
                   area = .data$area_m2,
                   area_diff = .data$area_diff,
                   vol = .data$vol_m3,
                   vol_diff = .data$vol_diff,
                   mean_depth = .data$mean_depth,
                   mean_diff = .data$mean_diff,
                   max_depth = .data$max_depth,
                   max_diff = .data$max_diff)

# Convert to English units
table_df <- table_df %>%
            mutate_at(c("elev", "elev_diff", "mean_depth", "mean_diff", "max_depth", "max_diff"),
                      NISTmeterTOft) %>%
            mutate_at(c("area", "area_diff"), NISTsqrMeterTOacre) %>%
            mutate_at(c("vol", "vol_diff"), NISTcubMeterTOacreFt)

# Round appropriately
round_precision <- 0.1
for (col in c("elev", "elev_diff", "mean_depth", "mean_diff", "max_depth", 
              "max_diff", "area", "area_diff")) {
  table_df[, col] <- round((1/round_precision)*table_df[, col])/(1/round_precision)

}
round_precision <- 1
for (col in c("vol", "vol_diff")) {
  table_df[, col] <- round((1/round_precision)*table_df[, col])/(1/round_precision)

}

# Convert to string
table_df <- table_df %>%
            mutate(elev = sprintf("%s \U00B1 %s", .data$elev, .data$elev_diff),
                   area = sprintf("%s \U00B1 %s", .data$area, .data$area_diff),
                   vol = sprintf("%s \U00B1 %s", .data$vol, .data$vol_diff),
                   mean_depth = sprintf("%s \U00B1 %s", .data$mean_depth, .data$mean_diff),
                   max_depth = sprintf("%s \U00B1 %s", .data$max_depth, .data$max_diff)) %>%
            select(.data$lake, .data$variable, .data$elev, .data$area, 
                   .data$vol, .data$mean_depth, .data$max_depth)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

### Exceedance Probability Ranges
```{r exceedance_range_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "exceedance_range"
metric_title    <- "**Exceedance Probability Range (ft)**"
variable_breaks <- c("a_10_90","a_25_75","m_10_90","m_25_75")
variable_labels <- c("Annual 10%-90%", "Annual 25%-75%", 
                     "Monthly 10%-90%", "Monthly 25%-75%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

## 2. Frequency
### Exceedances per decade (monthly timeseries)
```{r num_exceedances_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "num_dur_decade"
metric_title    <- "**Exceedances per Decade**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Exceedances per decade (seasonal timeseries)
```{r num_exceedances_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "num_dur_decade_seas"
metric_title    <- "**Exceedances per Decade**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Number of 2+ yr exceedances (monthly timeseries)
```{r num_2yr_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "num_2yr"
metric_title    <- "**Number of 2yr+ Exceedances**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Number of 2+ yr exceedances (seasonal timeseries)
```{r num_2yr_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "num_2yr_seas"
metric_title    <- "**Number of 2yr+ Exceedances**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Number of 2+ yr exceedances per decade (monthly timeseries)
```{r num_2yr_decade_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "num_2yr_decade"
metric_title    <- "**Number of 2yr+ Exceedances Per Decade**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Number of 2+ yr exceedances per decade (seasonal timeseries)
```{r num_2yr_decade_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "num_2yr_decade_seas"
metric_title    <- "**Number of 2yr+ Exceedances Per Decade**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Recurrence Interval of 2+ yr exceedances (monthly timeseries)
```{r recurr_int_2yr_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "recurr_int_2yr"
metric_title    <- "**Recurrence Interval (yrs) for 2yr+ exceedances**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Recurrence Interval of 2+ yr exceedances (seasonal timeseries)
```{r recurr_int_2yr_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "recurr_int_2yr_seas"
metric_title    <- "**Reccurrance Interval (yrs) for 2yr+ exceedances**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Frequency of 1ft departure from median
```{r depart_median_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "depart_median"
metric_title    <- "**Probability within 1 ft of median (%)**"
variable_breaks <- c("0.3m", "-0.3m")
variable_labels <- c("+1ft", "-1ft")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

## 3. Duration
### Duration histogram (representative timeseries)
```{r duration, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=7}
plot_duration(hist_levels_short, 
              probs = c(10, 25, 50, 75, 90), 
              show_median = TRUE, 
              bin_size = 1, 
              text_size = text_size,
              max_months = 12)
```

<br>

### Median duration (monthly timeseries)
```{r median_dur_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_dur"
metric_title    <- "**Median Duration (months)**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Median duration (seasonal timeseries)
```{r median_dur_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_dur_seas"
metric_title    <- "**Median Duration (months)**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### CV of duration (monthly timeseries)
```{r cv_dur_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_dur"
metric_title    <- "**CV of Duration (%)**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### CV of duration (seasonal timeseries)
```{r cv_dur_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_dur_seas"
metric_title    <- "**CV of Duration (%)**"
variable_breaks <- c("10","25","a50_1","a50","b50","b50_1","75","90")
variable_labels <- c("10%","25%","50%+1ft",">50%","<50%","50%-1ft","75%","90%")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

## 4. Rate of change
### Rate of change histogram (representative timeseries)
```{r rate_of_change, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=7}
plot_rate_change(hist_levels_short, text_size = text_size)
```
<br>

### Median rise rate
```{r median_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_rise_rate"
metric_title    <- "**Median Rise Rate (ft/time period)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### CV rise rate
```{r cv_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_rise_rate"
metric_title    <- "**CV of Rise Rate (%)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Number of fast rise rates (times)
```{r fast_rise_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "fast_rise"
metric_title    <- "**Number of Fast Rise Rates (number of months)**"
variable_breaks <- c("rate_1_5ft", "rate_3ft")
variable_labels <- c("12 mo. change > 1.5 ft", "12 mo. change > 3 ft")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

### Frequency of fast rise rates (times per decade)
```{r fast_rise_rate_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "fast_rise_decade"
metric_title    <- "**Frequency of Fast Rise Rates (months per decade)**"
variable_breaks <- c("rate_1_5ft", "rate_3ft")
variable_labels <- c("12 mo. change > 1.5 ft", "12 mo. change > 3 ft")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Median fall rate
```{r median_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "median_fall_rate"
metric_title    <- "**Median Fall Rate (ft/time period)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "meterTOft",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### CV fall rate
```{r cv_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "cv_fall_rate"
metric_title    <- "**CV of Fall Rate (%)**"
variable_breaks <- "sort"
variable_labels <- "breaks"

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

### Number of fast fall rates (times)
```{r fast_fall_rate_01, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "fast_fall"
metric_title    <- "**Number of Fast Fall Rates (number of months)**"
variable_breaks <- c("rate_1_5ft", "rate_3ft")
variable_labels <- c("12 mo. change < -1.5 ft", "12 mo. change < -3 ft")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

<br>

### Frequency of fast fall rates (times per decade)
```{r fast_fall_rate_02, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "fast_fall_decade"
metric_title    <- "**Frequency of Fast Fall Rates (months per decade)**"
variable_breaks <- c("rate_1_5ft", "rate_3ft")
variable_labels <- c("12 mo. change < -1.5 ft", "12 mo. change < -3 ft")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```
<br>

## 5. Timing

### Good Spawning Years
```{r good_spawning, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=6.5, fig.height=6}
metric_name     <- "good_spawning"
metric_title    <- "**Good Spawning Years (%)**"
variable_breaks <- c("high_spring", "steady_summer", "good_spawning")
variable_labels <- c("High April Levels", "Steady Summer Levels", "Good Spawning Year")

table_df <- create_table(hist_metrics_short, metric_name, variable_breaks, 
                         variable_labels, convert_units = "",
                         round_precision = 0.1)

kableExtra::kable(table_df, caption = metric_title)
```

### Annual vs. Monthly ranges

```{r timing, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.width=4, fig.height=3}
plot_df <- hist_metrics_short %>%
           filter(.data$metric == "exceedance_range") %>%
           mutate(time = str_sub(.data$variable, 1, 1),
                  range = str_sub(.data$variable, 3, -1),
                  median = NISTunits::NISTmeterTOft(.data$median),
                  sd = NISTunits::NISTmeterTOft(.data$sd)) %>%
           mutate(range = str_replace(.data$range, "10_90", "10%-90%")) %>%
           mutate(range = str_replace(.data$range, "25_75", "25%-75%")) %>%
           mutate(text = sprintf("%s \U00B1 %s", .data$median, .data$sd)) %>%
           # filter(.data$range != "25%-75%") %>%
           select(.data$lake, .data$time, .data$range, .data$median, .data$sd)
plot_df <- dcast(plot_df, lake+range~time, value.var = "median")


ggplot(plot_df) +
  geom_point(aes(x = a, y = m, shape = lake, color = range), 
             # color = "black",
             size = 2) +
  geom_abline(slope = 1, intercept = 0) +
  labs(x = "Annual Range (ft)",
       y = "Monthly Range (ft)",
       shape = "") +
  theme_bw() +
  scale_color_manual(name = "",
                     breaks = c("10%-90%", "25%-75%"),
                     values = c("black", "grey70")) +
  theme(text = element_text(family = "Segoe UI Semilight",
                            size = text_size),
        plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        legend.box = "vertical",
        legend.margin = margin(0,0,0,0))
```
